---
title: "Wildlife analysis"
author: "Stanley"
date: "23/07/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggthemes)
wildlife <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-23/wildlife_impacts.csv")
View(wildlife)
```


# Attempt at data visualization 

No indepth analysis is carried out, just plain basic data visualization is documented here. This script would like to attempt the following question:

- which species have the highest reported kills?
- How does bird kills evolved over time?
- At what time of the day is the kills mostly occuring?

Before we begin just a minor one-liner EDA.

```{r echo=FALSE}
wildlife %>% 
  count(species) %>% 
  arrange(desc(n))


wildlife %>% 
  count(species) %>% 
  filter(!str_detect(species, "Unknown")) %>% 
  summarise(total = mean(n))
```


So let begin answering some of those question, shall we!

## which species have the highest reported kills?


```{r}
highest_bird_kill <- wildlife %>% 
  count(species) %>% 
  filter(!str_detect(species, "Unknown")) %>% 
  arrange(desc(n)) %>% 
  head(20) %>% 
  mutate(species = fct_reorder(species, n)) %>% 
  ggplot(aes(species, n)) +
  geom_col() +
  coord_flip() +
  geom_hline(yintercept = 750, linetype = "dashed") +
  labs(x = "Bird species", 
       y = "Number of kills", 
       title = "Birds with the highest kills", 
       subtitle = "Approximately 80% of Gulls and Mourning dove are kills reported by airline operators", 
       caption = "Source: #Tidytueday submit | Data viz: @nmorstanlee") +
  theme_solarized()

ggsave("highest_bird_kill.png", highest_bird_kill, dpi = 400)
```



# Annual evolution of the kills?

```{r}
annual_bird_kill <- wildlife %>% 
  count(species, year = incident_year) %>% 
  filter(!str_detect(species, "Unknown")) %>% 
  group_by(year) %>% 
  summarise(n = sum(n)) %>% 
  ungroup() %>% 
  ggplot(aes(year, n)) +
  geom_line() +
  geom_rect(data = data.frame(x1 = 2008, x2 = Inf, y1 = 0, y2 = Inf), 
            aes(xmin = x1, xmax = x2, ymin = y1, ymax = y2), inherit.aes = FALSE, fill = "pink", alpha = 0.3) +
  geom_vline(xintercept = c(2008, Inf), linetype = "dashed") +
  scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2018), 
                     labels = c(1990, 1995, 2000, 2005, 2010, 2015, 2018)) +
   scale_y_continuous(breaks = c(500, 1000, 1500, 2000, 2500, 3000), 
                     labels = c(500, 1000, 1500, 2000, 2500, 3000)) +

      labs(x = "Year",
       y = "Number of bird kills", 
       title = "Increase in incident of Bird Mortality since 2008", 
       subtitle = "Pink areas indicate zone of increased avian death as seen by airline operators", 
       caption = "Source: #Tidytueday submit | Data viz: @nmorstanlee") +
  theme_economist()

ggsave("annual_bird_kill.png", annual_bird_kill, dpi = 400)


```


## Which time of the day is the kills mostly occuring?

```{r}
time_occurence_of_bird_kill <- wildlife %>% 
  count(species, time_of_day) %>% 
  group_by(time_of_day) %>% 
  summarise(n = sum(n)) %>% 
  filter(!is.na(time_of_day)) %>% 
  ggplot(aes(time_of_day, n)) +
  geom_col() +
  labs(x = "Time of day", 
       y = "Total kills per time of day", 
       title = "Which time of the day is bird mortality prominent", 
       subtitle = "Large number of birds are kills in broad-day light!!", 
       caption = "Source: #Tidytueday submit | Data viz: @nmorstanlee") +
  theme_solarized()

ggsave("time_occurence_of_bird_kill.png", time_occurence_of_bird_kill, dpi = 400)
```


## Miscellaneous graphs

```{r echo=FALSE, eval=FALSE, include=FALSE}
wildlife %>% 
  count(species, incident_month) %>% 
  group_by(incident_month) %>% 
  summarise(n = sum(n)) %>% 
  mutate(incident_month = month.name[incident_month], 
         incident_month = fct_reorder(incident_month, n)) %>% 
  ggplot(aes(incident_month, n)) +
  geom_col()
```


which state is the occurence of the kills most?
```{r echo=FALSE, eval=FALSE, include=FALSE}
wildlife %>% 
  filter(state %in% state.abb) %>% 
  count(state)
```
